### 2024-11-14 paper/paper.qmd
Add files via upload

---
title: "Assessing Music Trends of Beyoncé and Coldplay"
subtitle: "Song Duration and Valence Analysis Using Spotify Data"
author: 
  - Andy Jiang
  - Onon Burentuvshin
  - Rui Hu
thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/spotify)."
date: today
date-format: long
abstract: "This paper analyzes the audio features of songs from two globally renowned artists, Beyoncé and Coldplay, using Spotify's API. By focusing on song duration and valence (musical positivity), the study compares trends in their music over time. The results show that Beyoncé's songs vary more in length and tend to have higher valence, while Coldplay's tracks are more consistent in duration but exhibit a more melancholic tone."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(spotifyr)
library(lubridate)
library(ggplot2)

coldplay <- read_rds(here::here("data/01-raw_data/coldplay.rds"))
beyonce <- read_rds(here::here("data/01-raw_data/beyonce.rds"))
```

# Introduction

In recent years, music streaming platforms have provided vast amounts of data that enable a deeper analysis of the characteristics of different artists’ music. By utilizing Spotify’s audio feature data, we can examine several aspects of a song such as duration, tempo, energy, and valence (a measure of musical positivity or happiness). In this analysis, we focus on two globally renowned artists, Beyoncé and Coldplay, to compare key audio features of their music over time.

This study examines the duration and valence of songs released by Beyoncé and Coldplay, offering insights into how these attributes vary between the two artists. The analysis is visualized through boxplots and scatter plots, helping to reveal trends in their musical style and production across different time periods. The results suggest that while Beyoncé’s music varies more in duration and has a higher valence (indicating happier music), Coldplay’s songs are generally more consistent in length, with less valence or happiness expressed in their tracks.

# Data {#sec-data}
We use the statistical programming language R [@citeR], and library tidyverse [@tidyverse], spotifyr [@citeSpotifyr], lubridate [@lubridate], ggplot2 [@citeGG]. The data is from Spotify [@spotify].

## Overview
The key attributes of the data include:

Duration: Measured in milliseconds, this represents the length of each track.
Valence: A measure from 0 to 1 indicating how happy or sad a track sounds. Higher values correspond to happier, more cheerful music, while lower values indicate sadness or melancholy.
Album release date: The date when each album was released, which helps in analyzing trends over time.
Using this dataset, we compared the two artists on two main dimensions: song duration and valence.

@sec-results discusses our findings on the song duration and valence between the two artists.

# Results {#sec-results}
Our results on song duration are summarized in @fig-duration-beyonce and @fig-duration-coldplay. And our results on valence are summerized in @fig-valence.

```{r}
#| label: fig-duration-beyonce
#| fig-cap: Duration of Songs of Beyonce
#| echo: false
#| warning: false
#| message: false

beyonce <- as_tibble(beyonce)

beyonce |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

```{r}
#| label: fig-duration-coldplay
#| fig-cap: Duration of Songs of Coldplay
#| echo: false
#| warning: false
#| message: false

coldplay <- as_tibble(coldplay)

coldplay |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

```{r}
#| label: fig-valence
#| fig-cap: Valence of the Two Artists' Songs
#| echo: false
#| warning: false
#| message: false

rbind(beyonce, coldplay) |>
  select(artist_name, album_release_date, valence) |>
  mutate(album_release_date = ymd(album_release_date)) |> 
  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  theme_minimal() +
  facet_wrap(facets = vars(artist_name), dir = "v") +
  labs(
    x = "Album release date",
    y = "Valence",
    color = "Artist"
  ) +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

## Song Duration

@fig-duration-beyonce and @fig-duration-coldplay illustrate song duration for Beyoncé and Coldplay over time. Several interesting observations can be made from these visualizations:

Beyoncé: The song duration for Beyoncé shows a wider range across the years, indicating a greater variability in how long her songs tend to be. Some of her albums contain shorter tracks, while others include significantly longer ones. This suggests that Beyoncé experiments more with song length, perhaps catering to different album concepts or genres.

Coldplay: On the other hand, Coldplay’s songs tend to be more consistent in duration, with less variation in length. While there are a few outliers with exceptionally long or short tracks, most of Coldplay’s songs stay within a narrower duration range. This may reflect Coldplay’s commitment to a particular style or format for their music.

## Valence

@fig-valence shows the valence, or happiness, of songs by Beyoncé and Coldplay over time. This is where the two artists differ most notably:

Beyoncé: Her songs have generally higher valence scores, meaning that her music tends to sound more upbeat, positive, and emotionally vibrant. This aligns with the perception of Beyoncé’s music, which often conveys themes of empowerment, celebration, and joy.

Coldplay: Coldplay’s music, in contrast, exhibits lower valence scores, suggesting a more melancholic or introspective tone. Coldplay is known for producing emotionally resonant tracks, often focusing on themes of love, loss, and reflection, which could explain the lower valence values.

\newpage

# References

### 2024-11-14 scripts/02-download_data.R

@@ -1,0 +1,23 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from spotify
+# Author: Andy Jiang
+# Date: 10 October 2024
+# Contact: ayx.jiang@mail.utoronto.ca
+# Pre-requisites: None
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(spotifyr)
+library(readr)
+
+#### Download data ####
+radiohead <- get_artist_audio_features("radiohead")
+saveRDS(radiohead, "data/01-raw_data/radiohead.rds")
+
+coldplay <- get_artist_audio_features("coldplay")
+saveRDS(coldplay, "data/01-raw_data/coldplay.rds")
+
+beyonce <- get_artist_audio_features("beyonce")
+saveRDS(beyonce, "data/01-raw_data/beyonce.rds")
+

### 2024-11-14 paper/paper.qmd
update script

@@
 ---
-title: "Assessing Music Trends of Beyoncé and Coldplay"
-subtitle: "Song Duration and Valence Analysis Using Spotify Data"
+title: ""
+subtitle: ""
 author: 
   - Andy Jiang
-  - Onon Burentuvshin
-  - Rui Hu
-thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/spotify)."
+thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/grocery-price-analysis.git)."
 date: today
 date-format: long
-abstract: "This paper analyzes the audio features of songs from two globally renowned artists, Beyoncé and Coldplay, using Spotify's API.
-By focusing on song duration and valence (musical positivity), the study compares trends in their music over time.
-The results show that Beyoncé's songs vary more in length and tend to have higher valence, while Coldplay's tracks are more consistent in duration but exhibit a more melancholic tone."
+abstract: ""
 format: pdf
@@
 library(tidyverse)
-library(spotifyr)
-library(lubridate)
+library(DBI)
+library(RSQLite)
 library(ggplot2)
 
-coldplay <- read_rds(here::here("data/01-raw_data/coldplay.rds"))
-beyonce <- read_rds(here::here("data/01-raw_data/beyonce.rds"))
+# Connect to the SQLite database
+con <- dbConnect(RSQLite::SQLite(), "/Users/y4nxunj/Downloads/hammer-2-processed.sqlite")  # Update path
+
+# Run the join query from '01_join_raw_and_product.sql' to get combined data
+query <- "
+  SELECT 
+      r.nowtime, 
+      p.vendor, 
+      r.product_id, 
+      p.product_name, 
+      p.brand, 
+      r.current_price, 
+      r.old_price, 
+      p.units, 
+      r.price_per_unit, 
+      r.other 
+  FROM 
+      raw AS r
+  INNER JOIN 
+      product AS p ON r.product_id = p.id
+  ORDER BY 
+      r.nowtime;
+"
+data <- dbGetQuery(con, query)
+
+# Disconnect from the database
+dbDisconnect(con)
+
+# Now you have 'data' in R and can use it for visualizations, e.g., average price by vendor
+avg_price_by_vendor <- aggregate(current_price ~ vendor, data, mean)
+
+ggplot(avg_price_by_vendor, aes(x = vendor, y = current_price, fill = vendor)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Average Price by Vendor",
+       x = "Vendor", y = "Average Price") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
 ```
@@
 ```{r}
-#| label: fig-duration-beyonce
-#| fig-cap: Duration of Songs of Beyonce
-#| echo: false
-#| warning: false
-#| message: false
-
-beyonce <- as_tibble(beyonce)
-
-beyonce |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(
-    x = album_release_date,
-    y = duration_ms,
-    group = album_release_date
-  )) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
-  theme_minimal() +
-  labs(
-    x = "Album release date",
-    y = "Duration of song (ms)"
-  )
-```
-
-```{r}
 #| label: fig-duration-coldplay
@@
 #| message: false
+# Filter for rows where products were on sale
+sales_data <- subset(data, !is.na(old_price))
 
-coldplay <- as_tibble(coldplay)
+# Count sales occurrences by vendor
+sales_count <- as.data.frame(table(sales_data$vendor))
+colnames(sales_count) <- c("vendor", "sale_count")
 
-coldplay |>
-  mutate(album_release_date = ymd(album_release_date)) |>
-  ggplot(aes(
-    x = album_release_date,
-    y = duration_ms,
-    group = album_release_date
-  )) +
-  geom_boxplot() +
-  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
+# Plot frequency of sales by vendor
+ggplot(sales_count, aes(x = vendor, y = sale_count, fill = vendor)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Frequency of Sales by Vendor",
+       x = "Vendor", y = "Sale Count") +
   theme_minimal() +
-  labs(
-    x = "Album release date",
-    y = "Duration of song (ms)"
-  )
+  theme(legend.position = "none")
+
 ```
@@
 
-rbind(beyonce, coldplay) |>
-  select(artist_name, album_release_date, valence) |>
-  mutate(album_release_date = ymd(album_release_date)) |> 
-  ggplot(aes( x = album_release_date, y = valence, color = artist_name)) +
-  geom_point(alpha = 0.3) +
-  geom_smooth() +
+# Improved box plot with fewer jitter points and adjusted transparency and size
+set.seed(42)  # For reproducible sampling
+sampled_data <- data[sample(nrow(data), size = min(5000, nrow(data))), ]  # Sample 5000 points or fewer
+
+ggplot(data, aes(x = vendor, y = current_price, fill = vendor)) +
+  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Hide outliers in box plot
+  geom_jitter(data = sampled_data, width = 0.2, alpha = 0.1, size = 0.5, color = "black") +  # Sampled points with low transparency
+  scale_y_continuous(trans = 'log10', limits = c(1, 100)) +  # Log scale and limit y-axis
+  labs(title = "Price Distribution Across Vendors",
+       x = "Vendor", y = "Current Price (Log Scale)") +
   theme_minimal() +
-  facet_wrap(facets = vars(artist_name), dir = "v") +
-  labs(
-    x = "Album release date",
-    y = "Valence",
-    color = "Artist"
-  ) +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
+  theme(legend.position = "none",
+        plot.title = element_text(size = 16, face = "bold"),
+        axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
+  scale_fill_brewer(palette = "Set3")  # Use a color palette for better presentation
+
+```
+
+```{r}
+# Plot distribution of prices by vendor
+# Calculate average price per vendor
+avg_price_by_vendor <- aggregate(current_price ~ vendor, data, mean, na.rm = TRUE)
+
+# Plot average price by vendor
+ggplot(avg_price_by_vendor, aes(x = vendor, y = current_price, fill = vendor)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Average Price by Vendor",
+       x = "Vendor", y = "Average Price") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
 ```

### 2024-11-14 scripts/02-download_data.R
update script

@@ -1,23 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from spotify
-# Author: Andy Jiang
-# Date: 10 October 2024
-# Contact: ayx.jiang@mail.utoronto.ca
-# Pre-requisites: None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(spotifyr)
-library(readr)
-
-#### Download data ####
-radiohead <- get_artist_audio_features("radiohead")
-saveRDS(radiohead, "data/01-raw_data/radiohead.rds")
-
-coldplay <- get_artist_audio_features("coldplay")
-saveRDS(coldplay, "data/01-raw_data/coldplay.rds")
-
-beyonce <- get_artist_audio_features("beyonce")
-saveRDS(beyonce, "data/01-raw_data/beyonce.rds")
-

### 2024-11-14 paper/paper.qmd
update paper.qmd and citation

@@
 ---
-title: ""
-subtitle: ""
+title: "Grocery Pricing Strategies: An Analysis of Vendor Positioning and Market Dynamics"
+subtitle: "Examining Sales Frequency, Price Distribution, and Average Price Levels Across Major Grocery Chains"
 author: 
@@
 date-format: long
-abstract: ""
+abstract: "This paper examines grocery pricing strategies across major vendors to understand how they position themselves in a competitive market.
+By analyzing sales frequency, price distribution, and average price levels, we identify distinct approaches used by each vendor to attract customers.
+Our findings reveal that some vendors, like Metro, rely on frequent promotions to appeal to a broad customer base, while others, such as No Frills and Walmart, prioritize everyday low pricing to cater to budget-conscious shoppers.
+This analysis highlights the varied tactics grocery vendors use to target different consumer segments, providing insights into the competitive dynamics that shape grocery pricing and customer choices."
 format: pdf
@@
 # Introduction
+In recent years, competitive pricing and promotional strategies among grocery vendors have created a rich landscape for data analysis.
+By examining grocery price data collected across multiple vendors, we can gain insights into various aspects of vendor pricing, such as frequency of sales, price distribution, and average price levels.
+This analysis focuses on several major grocery vendors, comparing their sales frequency, price ranges, and average prices to better understand their market positioning and strategies.
 
-In recent years, music streaming platforms have provided vast amounts of data that enable a deeper analysis of the characteristics of different artists’ music.
-By utilizing Spotify’s audio feature data, we can examine several aspects of a song such as duration, tempo, energy, and valence (a measure of musical positivity or happiness).
-In this analysis, we focus on two globally renowned artists, Beyoncé and Coldplay, to compare key audio features of their music over time.
-
-This study examines the duration and valence of songs released by Beyoncé and Coldplay, offering insights into how these attributes vary between the two artists.
-The analysis is visualized through boxplots and scatter plots, helping to reveal trends in their musical style and production across different time periods.
-The results suggest that while Beyoncé’s music varies more in duration and has a higher valence (indicating happier music), Coldplay’s songs are generally more consistent in length, with less valence or happiness expressed in their tracks.
+This study investigates the frequency of sales events, the distribution of product prices, and the average price by vendor, highlighting the distinct approaches taken by each vendor to attract and retain customers.
+Through visualizations such as bar plots and box plots, the analysis reveals trends in vendor pricing strategies, showcasing how some vendors prioritize frequent discounts, while others maintain a steady low-price approach.
+The findings suggest that vendors like Metro employ high-frequency sales with a broad price range, while No Frills and Walmart focus on consistently low prices.
+These patterns reflect the varied strategies used by grocery vendors to appeal to different consumer segments.
 
 # Data {#sec-data}
-We use the statistical programming language R [@citeR], and library tidyverse [@tidyverse], spotifyr [@citeSpotifyr], lubridate [@lubridate], ggplot2 [@citeGG]. The data is from Spotify [@spotify].
+We use the statistical programming language R [@citeR] and several libraries including tidyverse [@tidyverse], DBI [@citeDBI], RSQLite [@citeRSQLite], and ggplot2 [@citeGG].
+The data was sourced from a project database created by Jacob Filipp, containing grocery price data collected across various vendors.
 
@@
 
-Duration: Measured in milliseconds, this represents the length of each track.
-Valence: A measure from 0 to 1 indicating how happy or sad a track sounds. Higher values correspond to happier, more cheerful music, while lower values indicate sadness or melancholy.
-Album release date: The date when each album was released, which helps in analyzing trends over time.
-Using this dataset, we compared the two artists on two main dimensions: song duration and valence.
+Current Price: The price of each product at the time of data collection, which helps in understanding price distribution across vendors.
+Sales Frequency: The frequency of discounted prices, indicating how often each vendor holds sales or promotional events.
+Vendor: The specific grocery store or vendor from which the pricing data was collected, allowing for comparisons between different grocery chains.
+Using this dataset, we analyzed the pricing and promotional strategies of each vendor, focusing on three main dimensions: frequency of sales, price distribution, and average price by vendor.
 
-@sec-results discusses our findings on the song duration and valence between the two artists.
+@sec-results presents our findings on these aspects and highlights the distinct strategies employed by different vendors.
 
 # Results {#sec-results}
-Our results on song duration are summarized in @fig-duration-beyonce and @fig-duration-coldplay. And our results on valence are summerized in @fig-valence.
+Our results on the frequency of sales by vendor is illustrated in @fig-sales-frequency, while @fig-price-distribution shows the distribution of prices across vendors.
+Additionally, @fig-average-price highlights the average price by vendor.
 
 ```{r}
-#| label: fig-duration-coldplay
-#| fig-cap: Duration of Songs of Coldplay
+#| label: fig-sales-frequency
+#| fig-cap: The Frequency of Sales by Vendor
 #| echo: false
@@
 ```{r}
-#| label: fig-valence
-#| fig-cap: Valence of the Two Artists' Songs
+#| label: fig-price-distribution
+#| fig-cap: The Price Distribution Across Vendors
 #| echo: false
@@
 
-# Improved box plot with fewer jitter points and adjusted transparency and size
-set.seed(42)  # For reproducible sampling
-sampled_data <- data[sample(nrow(data), size = min(5000, nrow(data))), ]  # Sample 5000 points or fewer
+# Ensure current_price is numeric and handle any non-numeric values
+data$current_price <- as.numeric(data$current_price)
 
-ggplot(data, aes(x = vendor, y = current_price, fill = vendor)) +
-  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Hide outliers in box plot
-  geom_jitter(data = sampled_data, width = 0.2, alpha = 0.1, size = 0.5, color = "black") +  # Sampled points with low transparency
-  scale_y_continuous(trans = 'log10', limits = c(1, 100)) +  # Log scale and limit y-axis
+# Remove rows with NA, zero, or negative values in current_price
+plot_data <- data %>%
+  filter(!is.na(current_price) & current_price > 0)
+
+# Set seed and sample data
+set.seed(42)
+sampled_data <- plot_data[sample(seq_len(nrow(plot_data)), size = min(5000, nrow(plot_data))), ]
+
+# Create box plot
+ggplot(plot_data, aes(x = vendor, y = current_price, fill = vendor)) +
+  geom_boxplot(outlier.shape = NA, alpha = 0.6) +
+  geom_jitter(data = sampled_data, width = 0.2, alpha = 0.1, size = 0.5, color = "black") +
+  scale_y_continuous(trans = 'log10', limits = c(1, 100)) +
   labs(title = "Price Distribution Across Vendors",
@@
   theme_minimal() +
-  theme(legend.position = "none",
-        plot.title = element_text(size = 16, face = "bold"),
-        axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels for readability
-  scale_fill_brewer(palette = "Set3")  # Use a color palette for better presentation
+  theme(
+    legend.position = "none",
+    plot.title = element_text(size = 16),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  ) +
+  scale_fill_brewer(palette = "Set3")
 
@@
 ```{r}
+#| label: fig-average-price
+#| fig-cap: The Average Price by Vendor
+#| echo: false
+#| warning: false
+#| message: false
+
 # Plot distribution of prices by vendor
@@
 
-## Song Duration
+## Frequency of Sales by Vendor
+In @fig-sales-frequency, Metro stands out with a noticeably higher frequency of sales compared to all other vendors.
+This suggests that Metro might be employing an aggressive pricing strategy, using frequent sales promotions to attract a large customer base.
+In contrast, Loblaws and Voila follow at a second tier, with sales frequencies approximately half that of Metro.
+This moderate level of sales activity indicates a competitive, but less aggressive, approach.
+No Frills, Save-On-Foods, and Walmart have similar and relatively low frequencies, at around half the level of Loblaws and Voila.
+This low frequency suggests that these vendors may focus less on frequent promotional events and more on maintaining stable pricing or lower base prices.
 
-@fig-duration-beyonce and @fig-duration-coldplay illustrate song duration for Beyoncé and Coldplay over time. Several interesting observations can be made from these visualizations:
+## Price Distribution Across Vendors
 
-Beyoncé: The song duration for Beyoncé shows a wider range across the years, indicating a greater variability in how long her songs tend to be.
-Some of her albums contain shorter tracks, while others include significantly longer ones.
-This suggests that Beyoncé experiments more with song length, perhaps catering to different album concepts or genres.
+@fig-price-distribution reveals further distinctions in pricing strategy and positioning among the vendors.
+No Frills and Walmart show consistently lower price distributions, aligning with a budget-friendly image.
+This pattern suggests that these vendors aim to attract price-sensitive customers by offering lower-cost items, with minimal overlap into higher price ranges.
+Metro, on the other hand, displays a much wider price range, with a significant number of higher price points that are not seen in other vendors.
+This broad range suggests that Metro caters to a wide spectrum of customers, from budget-conscious shoppers to those willing to pay for premium products.
+Other vendors, such as Loblaws and Voila, show moderate price distributions, which are less spread out than Metro’s but still cover a range that suggests a balanced approach rather than strictly budget or premium positioning.
 
-Coldplay: On the other hand, Coldplay’s songs tend to be more consistent in duration, with less variation in length.
-While there are a few outliers with exceptionally long or short tracks, most of Coldplay’s songs stay within a narrower duration range.
-This may reflect Coldplay’s commitment to a particular style or format for their music.
+## Average Price by Vendor
+@fig-average-price aligns with the previous observations, highlighting Metro’s position as the highest-priced vendor on average.
+This high average price, combined with Metro’s wide distribution range, reinforces the notion that Metro offers a variety of products, including higher-priced or premium items.
+Galleria follows Metro with a relatively high average price, indicating a potential focus on specialty or premium goods.
+No Frills and Walmart emerge as the lowest-priced vendors on average, solidifying their positioning as budget-friendly options.
+The other vendors, such as Loblaws, Save-On-Foods, T&T, and Voila, occupy a middle ground, with average prices slightly above the lowest-priced vendors and below Galleria, suggesting a balanced pricing approach that caters to both cost-conscious and mid-range consumers.
 
-## Valence
+# Discussion {#sec-discussion}
+The results of this analysis reveal distinct pricing and promotional strategies among the grocery vendors, indicating how each vendor may be positioning itself within the market.
+Metro’s high frequency of sales, shown in the sales frequency graph, suggests a competitive pricing strategy that relies heavily on promotional events to attract customers.
+This approach could be aimed at creating a perception of savings, encouraging frequent purchases, and catering to a broad range of customers who respond to discounts.
+The presence of frequent sales might also reflect Metro’s intention to balance a diverse range of price points, as seen in the wide distribution of prices across products.
+This flexibility allows Metro to cater to both budget-conscious consumers and those willing to pay premium prices.
 
-@fig-valence shows the valence, or happiness, of songs by Beyoncé and Coldplay over time. This is where the two artists differ most notably:
+On the other hand, No Frills and Walmart demonstrate a consistent commitment to low prices, with lower price distributions and the lowest average prices among the vendors.
+Their relatively low frequency of sales suggests a focus on everyday low pricing rather than relying on promotions to drive customer traffic.
+This approach is likely attractive to price-sensitive shoppers who value predictably low prices over the fluctuations associated with frequent sales.
+The data implies that No Frills and Walmart have established themselves as budget-friendly options, strategically positioning themselves as reliable choices for consumers seeking affordability.
 
-Beyoncé: Her songs have generally higher valence scores, meaning that her music tends to sound more upbeat, positive, and emotionally vibrant.
-This aligns with the perception of Beyoncé’s music, which often conveys themes of empowerment, celebration, and joy.
+Galleria, with a high average price and fewer sales, appears to target a different segment.
+The relatively high prices suggest a focus on specialty or premium products, positioning Galleria as a niche or premium vendor within the market.
+This positioning may attract consumers who prioritize quality or unique offerings over low prices.
+The lack of high-frequency sales events implies a strategy that focuses less on discount-driven customer acquisition and more on brand or product differentiation, possibly appealing to a specific audience that values specialty goods.
 
-Coldplay: Coldplay’s music, in contrast, exhibits lower valence scores, suggesting a more melancholic or introspective tone.
-Coldplay is known for producing emotionally resonant tracks, often focusing on themes of love, loss, and reflection, which could explain the lower valence values.
+The remaining vendors, including Loblaws, Save-On-Foods, T&T, and Voila, demonstrate a more balanced approach.
+These vendors have moderate frequencies of sales and average prices, positioning them between the budget-oriented and premium segments.
+This strategy likely appeals to a broad consumer base that includes both cost-conscious shoppers and those willing to spend slightly more for a variety of options.
+By maintaining average prices that are neither the lowest nor the highest, these vendors may capture customers seeking value without the extremes of budget or premium positioning.
+
+Overall, the data illustrates a competitive landscape where each vendor’s pricing and promotional strategy plays a distinct role in shaping customer perception and loyalty.
+Metro’s aggressive promotional tactics, combined with a wide range of prices, allow it to attract a diverse audience.
+Meanwhile, No Frills and Walmart stand out as dependable low-cost options, appealing to value-driven customers.
+Galleria’s premium positioning caters to a niche market segment, and the remaining vendors capture a middle ground, appealing to customers looking for a balance of price and quality.
+These findings underscore the varied strategies used by grocery vendors to differentiate themselves and compete for different consumer segments in a competitive market.
 

### 2024-11-14 paper/paper.qmd
update paper, llm, readme

@@
 
+This paper is organized as follows: @sec-data outlines the dataset and its measurement of real-world pricing strategies.
+@sec-results presents key findings on sales frequency, price distribution, and average price, revealing distinct vendor pricing strategies.
+@sec-discussion addresses considerations such as correlation versus causation, missing data, and sources of bias, which are essential for interpreting the analysis accurately.
+
 # Data {#sec-data}
-We use the statistical programming language R [@citeR] and several libraries including tidyverse [@tidyverse], DBI [@citeDBI], RSQLite [@citeRSQLite], and ggplot2 [@citeGG].
-The data was sourced from a project database created by Jacob Filipp, containing grocery price data collected across various vendors.
+We use the statistical programming language R [@citeR] and several libraries, including tidyverse [@tidyverse], DBI [@citeDBI], RSQLite [@citeRSQLite], and ggplot2 [@citeGG].
+The data was sourced from Project Hammer [@Filipp_2024], a project database created by Jacob Filipp that compiles grocery price data from various vendors.
+
+## Measurement
+The dataset used in this analysis captures real-world grocery pricing and promotional strategies from multiple vendors.
+Each entry, sourced from Project Hammer [@Filipp_2024], represents a product listing, with variables such as current_price, old_price, vendor, and nowtime.
+These fields translate vendor pricing strategies into data points: current_price reflects the active selling price, while old_price denotes any discount or promotional pricing.
+The vendor field identifies the grocery chain, and nowtime timestamps each entry, allowing us to analyze trends over time.
+This structure enables an organized examination of pricing tactics across vendors by converting complex market behaviors into measurable data.
 
@@
 # Discussion {#sec-discussion}
-The results of this analysis reveal distinct pricing and promotional strategies among the grocery vendors, indicating how each vendor may be positioning itself within the market.
-Metro’s high frequency of sales, shown in the sales frequency graph, suggests a competitive pricing strategy that relies heavily on promotional events to attract customers.
-This approach could be aimed at creating a perception of savings, encouraging frequent purchases, and catering to a broad range of customers who respond to discounts.
-The presence of frequent sales might also reflect Metro’s intention to balance a diverse range of price points, as seen in the wide distribution of prices across products.
-This flexibility allows Metro to cater to both budget-conscious consumers and those willing to pay premium prices.
+The results of this analysis offer valuable insights into the pricing and promotional strategies of various grocery vendors.
+However, several methodological considerations need to be addressed to ensure an accurate interpretation of the data.
+Understanding the distinction between correlation and causation, handling missing data, and accounting for potential sources of bias are essential to refining the conclusions drawn from this study.
+Each of these factors is discussed in the following subsections.
 
-On the other hand, No Frills and Walmart demonstrate a consistent commitment to low prices, with lower price distributions and the lowest average prices among the vendors.
-Their relatively low frequency of sales suggests a focus on everyday low pricing rather than relying on promotions to drive customer traffic.
-This approach is likely attractive to price-sensitive shoppers who value predictably low prices over the fluctuations associated with frequent sales.
-The data implies that No Frills and Walmart have established themselves as budget-friendly options, strategically positioning themselves as reliable choices for consumers seeking affordability.
+## Correlation vs. Causation
+In analyzing grocery pricing and promotional strategies, it’s important to distinguish between correlation and causation.
+While our data reveals correlations—such as a relationship between high-frequency sales and lower average prices in certain vendors—these associations do not imply causation.
+For example, Metro’s high sales frequency correlates with its broader price range, but this does not necessarily mean that frequent sales cause price diversity, or vice versa.
+Factors like target market segmentation, inventory management, and strategic pricing decisions can all influence these metrics independently.
+Therefore, while the analysis highlights patterns in how different vendors structure pricing and promotions, it does not establish causative relationships between these variables.
 
-Galleria, with a high average price and fewer sales, appears to target a different segment.
-The relatively high prices suggest a focus on specialty or premium products, positioning Galleria as a niche or premium vendor within the market.
-This positioning may attract consumers who prioritize quality or unique offerings over low prices.
-The lack of high-frequency sales events implies a strategy that focuses less on discount-driven customer acquisition and more on brand or product differentiation, possibly appealing to a specific audience that values specialty goods.
+## Missing Data
+The dataset may contain missing values, particularly in fields such as current and old prices.
+These missing values can affect the accuracy of the analysis, especially in sections where price distribution and averages are calculated.
+For example, if a vendor has missing values during high promotional periods, their frequency of sales may be underestimated.
+Similarly, missing prices for premium products could bias the average price calculation downward.
+To mitigate this, we addressed missing data by filtering out incomplete rows for critical analyses; however, the absence of data for certain time periods or products still limits the representativeness of our findings.
 
-The remaining vendors, including Loblaws, Save-On-Foods, T&T, and Voila, demonstrate a more balanced approach.
-These vendors have moderate frequencies of sales and average prices, positioning them between the budget-oriented and premium segments.
-This strategy likely appeals to a broad consumer base that includes both cost-conscious shoppers and those willing to spend slightly more for a variety of options.
-By maintaining average prices that are neither the lowest nor the highest, these vendors may capture customers seeking value without the extremes of budget or premium positioning.
-
-Overall, the data illustrates a competitive landscape where each vendor’s pricing and promotional strategy plays a distinct role in shaping customer perception and loyalty.
-Metro’s aggressive promotional tactics, combined with a wide range of prices, allow it to attract a diverse audience.
-Meanwhile, No Frills and Walmart stand out as dependable low-cost options, appealing to value-driven customers.
-Galleria’s premium positioning caters to a niche market segment, and the remaining vendors capture a middle ground, appealing to customers looking for a balance of price and quality.
-These findings underscore the varied strategies used by grocery vendors to differentiate themselves and compete for different consumer segments in a competitive market.
+## Sources of Bias
+Several potential sources of bias could affect the analysis and interpretation of grocery pricing data.
+First, the data is derived from a limited selection of vendors, which may not represent all grocery stores in the region, particularly smaller, independent stores.
+Second, the data is based on specific geographic regions and may not account for regional pricing variations.
+Finally, since the data was collected from website listings, it may not reflect in-store promotions or exclusive discounts for loyalty program members, thus potentially underestimating actual promotional frequencies.
+These sources of bias should be considered when interpreting the results, as they may influence conclusions about vendor strategies and market positioning.
 

### 2024-11-14 paper/paper.qmd
polish paper

@@
 bibliography: references.bib
+toc: true
 ---
@@
 # Data {#sec-data}
-We use the statistical programming language R [@citeR] and several libraries, including tidyverse [@tidyverse], DBI [@citeDBI], RSQLite [@citeRSQLite], and ggplot2 [@citeGG].
+We use the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `DBI` [@citeDBI], `RSQLite` [@citeRSQLite], and `ggplot2` [@citeGG].
 The data was sourced from Project Hammer [@Filipp_2024], a project database created by Jacob Filipp that compiles grocery price data from various vendors.
@@
 The dataset used in this analysis captures real-world grocery pricing and promotional strategies from multiple vendors.
-Each entry, sourced from Project Hammer [@Filipp_2024], represents a product listing, with variables such as current_price, old_price, vendor, and nowtime.
-These fields translate vendor pricing strategies into data points: current_price reflects the active selling price, while old_price denotes any discount or promotional pricing.
-The vendor field identifies the grocery chain, and nowtime timestamps each entry, allowing us to analyze trends over time.
+Each entry, sourced from Project Hammer [@Filipp_2024], represents a product listing, with variables such as `current_price`, `old_price`, `vendor`, and `nowtime`.
+These fields translate vendor pricing strategies into data points: `current_price` reflects the active selling price, while `old_price` denotes any discount or promotional pricing.
+The vendor field identifies the grocery chain, and `nowtime` timestamps each entry, allowing us to analyze trends over time.
 This structure enables an organized examination of pricing tactics across vendors by converting complex market behaviors into measurable data.
@@
 # Discussion {#sec-discussion}
-The results of this analysis offer valuable insights into the pricing and promotional strategies of various grocery vendors.
-However, several methodological considerations need to be addressed to ensure an accurate interpretation of the data.
-Understanding the distinction between correlation and causation, handling missing data, and accounting for potential sources of bias are essential to refining the conclusions drawn from this study.
-Each of these factors is discussed in the following subsections.
+The results of this analysis provide valuable insights into the pricing and promotional strategies of various grocery vendors, revealing distinct approaches across vendors like Metro, No Frills, Walmart, and others.
+However, to interpret these findings accurately, it is essential to address certain methodological considerations.
+These include distinguishing between correlation and causation, handling missing data, and recognizing potential sources of bias.
+These factors are critical to refining the conclusions drawn from the study and are discussed in detail below.
 
 ## Correlation vs. Causation
-In analyzing grocery pricing and promotional strategies, it’s important to distinguish between correlation and causation.
-While our data reveals correlations—such as a relationship between high-frequency sales and lower average prices in certain vendors—these associations do not imply causation.
-For example, Metro’s high sales frequency correlates with its broader price range, but this does not necessarily mean that frequent sales cause price diversity, or vice versa.
-Factors like target market segmentation, inventory management, and strategic pricing decisions can all influence these metrics independently.
-Therefore, while the analysis highlights patterns in how different vendors structure pricing and promotions, it does not establish causative relationships between these variables.
+Our analysis revealed interesting correlations, such as Metro’s high frequency of sales being associated with a broad price range, suggesting a dual approach that includes both budget and premium offerings.
+Additionally, No Frills and Walmart showed consistently low prices and lower sales frequency, which aligns with a strategy that focuses on everyday low pricing.
+However, while these correlations are clear, they do not imply causation.
+For instance, Metro’s high sales frequency and diverse price range do not necessarily mean that frequent sales lead to a broader price distribution or that a wider range causes more frequent promotions.
+Both metrics could be influenced by other factors, such as Metro’s target market, inventory diversity, or management’s pricing strategy, which may independently affect both the frequency of sales and price diversity.
+
+Similarly, vendors like No Frills and Walmart, with lower prices and fewer promotional events, might reflect a commitment to stable, low-cost offerings rather than suggesting that low prices cause a lower sales frequency.
+These observed patterns highlight potential correlations in vendor strategies but do not establish a causal link.
+Further research with controlled variables would be required to determine whether these strategies actively drive changes in other metrics or if they are simply co-occurring features of each vendor’s broader approach.
 
 ## Missing Data
-The dataset may contain missing values, particularly in fields such as current and old prices.
-These missing values can affect the accuracy of the analysis, especially in sections where price distribution and averages are calculated.
-For example, if a vendor has missing values during high promotional periods, their frequency of sales may be underestimated.
-Similarly, missing prices for premium products could bias the average price calculation downward.
-To mitigate this, we addressed missing data by filtering out incomplete rows for critical analyses; however, the absence of data for certain time periods or products still limits the representativeness of our findings.
+The dataset contained missing values in key fields, such as `current_price` and `old_price`, which affected parts of the analysis, particularly in calculating accurate price averages and sales frequency.
+For instance, missing price data during high promotional periods for certain vendors could lead to an underestimation of their true sales frequency.
+This is particularly relevant for vendors like Metro, where sales frequency was a significant observation.
+If promotional pricing data was incomplete, Metro’s high frequency of sales might appear inflated or underestimated, depending on when the data was missing.
+Similarly, missing data in the `current_price` field for premium items could cause an underrepresentation of higher prices in the dataset, potentially leading to a lower calculated average price than is accurate.
+This would especially impact vendors with premium offerings, like Galleria, where higher average prices were noted.
+
+To mitigate the effects of missing data, we filtered out incomplete rows in critical analyses, ensuring that results were based only on fully recorded entries.
+However, this approach cannot account for the representativeness of missing data over time and across vendors, and therefore, the conclusions drawn may be limited by these gaps.
+Addressing missing data remains essential for improving the accuracy of this type of analysis in future research.
 
 ## Sources of Bias
-Several potential sources of bias could affect the analysis and interpretation of grocery pricing data.
-First, the data is derived from a limited selection of vendors, which may not represent all grocery stores in the region, particularly smaller, independent stores.
-Second, the data is based on specific geographic regions and may not account for regional pricing variations.
-Finally, since the data was collected from website listings, it may not reflect in-store promotions or exclusive discounts for loyalty program members, thus potentially underestimating actual promotional frequencies.
-These sources of bias should be considered when interpreting the results, as they may influence conclusions about vendor strategies and market positioning.
+The dataset is subject to several potential biases that may affect the analysis and interpretation of grocery pricing data.
+First, the data includes only a select number of vendors, possibly excluding smaller, independent stores that may have different pricing and promotional strategies.
+This limitation could skew our understanding of average pricing, as larger vendors with more data points (e.g., Metro, Walmart) might dominate the analysis, potentially overshadowing strategies that smaller vendors might use.
+Additionally, the data is region-specific and does not account for regional variations in grocery pricing, meaning that the findings may not fully represent pricing trends across different areas.
+
+Another important source of bias stems from the method of data collection.
+The data was gathered from online listings, which may differ from in-store pricing and promotions.
+For instance, in-store promotions or discounts exclusive to loyalty program members would not be captured, potentially underestimating the frequency and diversity of promotions offered by each vendor.
+This could particularly affect vendors like No Frills and Walmart, which tend to have stable, low prices and may offer in-store promotions not reflected online.
+Consequently, this method of data collection might result in an incomplete picture of vendor strategies, especially for promotional events and discounts.
+
+These biases should be carefully considered when interpreting the results, as they may influence conclusions about vendor positioning and market strategy.
+While this analysis highlights general trends across major vendors, understanding the full scope of grocery pricing dynamics would require addressing these biases in future research, potentially by expanding the dataset to include more vendors, regions, and in-store data.
 

### 2024-11-14 paper/paper.qmd
update paper

@@
   - Andy Jiang
+  - Rui Hu
+  - Onon Burentuvshin
 thanks: "Code and data are available at: [https://github.com/AndyYanxunJiang/grocery-price-analysis.git)."

### 2024-11-14 paper/paper.qmd
finalize paper

@@
 This limitation could skew our understanding of average pricing, as larger vendors with more data points (e.g., Metro, Walmart) might dominate the analysis, potentially overshadowing strategies that smaller vendors might use.
-Additionally, the data is region-specific and does not account for regional variations in grocery pricing, meaning that the findings may not fully represent pricing trends across different areas.
+Additionally, the data is region-specific and does not account for regional variations in grocery pricing, meaning that the findings may not fully represent pricing trends across different areas outside of our target population of vendors within Toronto.
 
@@
 This could particularly affect vendors like No Frills and Walmart, which tend to have stable, low prices and may offer in-store promotions not reflected online.
-Consequently, this method of data collection might result in an incomplete picture of vendor strategies, especially for promotional events and discounts.
+Furthermore, this method of data collection raises the problem of measurement bias, pertaining to in-store pick-up prices.
+These pick-up prices may differ from those that are available at in-store location or online delivery.
+The non-random exclusion of such prices bias our estimates for average prices and trends thereof.
+For instance, consumers choosing in-store pickup often differ in preferences, price sensitivity, and shopping habits from those opting for delivery.
+Thus, this method may only partially capture vendor strategies, especially regarding promotions and discounts, leading to an incomplete view of pricing dynamics.
+
+The dataset has evolved to capture a broader spectrum of consumer habits by progressively including a wider variety of goods in the basket [@Filipp_2024].
+Initially focused on a smaller set of products, the basket later expanded, aiming to better represent the diversity of grocery consumption.
+However, this inconsistency in product range over time introduces a potential source of bias, particularly in the analysis of price distribution and trends.
+Including a larger variety of goods can shift the average prices and potentially alter the price distribution across vendors.
+For instance, if the initial basket emphasized staple goods with relatively stable prices, while later data incorporated specialty items with higher price variability, observed price trends might reflect changes in product range rather than actual market shifts.
+This inconsistency potentially introduces a bias that renders any perceived trends could stem from the evolution of the dataset itself rather than from underlying price changes.
+Adjusting for this by creating comparable baskets over time or normalizing based on product categories could help mitigate this bias and provide a clearer understanding of true price patterns.
 

### 2024-11-14 paper/paper.qmd
finalize paper

@@
 
+Our ability to establish a casual relationship in the data is limited due to lack of experimental design.
+Although a lack of experimental design does diminish our ability to infer causality, it only does so to a certain degree [@tellingstories].
+To further uncover the causal relationships between vendor positioning and its market dynamics, experiments such as a regression discontinuity design or difference in differences design can be employed.
+However, these designs require a careful consideration of the structure of our dataset and identification of potential biases.
+
 ## Missing Data

### 2024-11-21 paper/paper.qmd
update paper

@@
 # Connect to the SQLite database
-con <- dbConnect(RSQLite::SQLite(), "/Users/y4nxunj/Downloads/hammer-2-processed.sqlite")  # Update path
+con <- dbConnect(RSQLite::SQLite(), "../data/hammer-2-processed.sqlite")
 
@@
   geom_bar(stat = "identity") +
-  labs(title = "Frequency of Sales by Vendor",
-       x = "Vendor", y = "Sale Count") +
+  labs(x = "Vendor", y = "Sale Count") +
   theme_minimal() +
@@
   scale_y_continuous(trans = 'log10', limits = c(1, 100)) +
-  labs(title = "Price Distribution Across Vendors",
-       x = "Vendor", y = "Current Price (Log Scale)") +
+  labs(x = "Vendor", y = "Current Price (Log Scale)") +
   theme_minimal() +
@@
   geom_bar(stat = "identity") +
-  labs(title = "Average Price by Vendor",
-       x = "Vendor", y = "Average Price") +
+  labs(x = "Vendor", y = "Average Price") +
   theme_minimal() +

### 2024-11-21 scripts/06_save_data.R
simulate and test data

@@ -1,0 +1,9 @@
+source_path <- "/Users/y4nxunj/Downloads/hammer-2-processed.sqlite"
+destination_path <- "/Users/y4nxunj/Downloads/grocery-price-analysis/data/hammer-2-processed.sqlite"
+
+# Copy the file
+if (file.copy(source_path, destination_path)) {
+  message("File successfully copied to the data folder.")
+} else {
+  message("Failed to copy the file. Check paths and permissions.")
+}

### 2024-11-21 scripts/08_test_data.R

@@ -1,0 +1,84 @@
+library(DBI)
+library(RSQLite)
+
+# Create a temporary SQLite database
+con <- dbConnect(SQLite(), ":memory:")
+
+# Step 1: Create the 'product' table
+dbExecute(con, "
+CREATE TABLE product (
+    id INTEGER PRIMARY KEY,
+    product_name TEXT,
+    brand TEXT,
+    vendor TEXT
+);
+")
+
+# Insert data into the 'product' table
+dbExecute(con, "
+INSERT INTO product (id, product_name, brand, vendor) VALUES
+(1, 'Product 1', 'Brand X', 'Vendor A'),
+(2, 'Product 2', 'Brand Y', 'Vendor B'),
+(3, 'Product 3', 'Brand Z', 'Vendor C'),
+(4, 'Product 4', 'Brand X', 'Vendor D'),
+(5, 'Product 5', 'Brand Y', 'Vendor A'),
+(6, 'Product 6', 'Brand Z', 'Vendor B'),
+(7, 'Product 7', 'Brand X', 'Vendor C'),
+(8, 'Product 8', 'Brand Y', 'Vendor D'),
+(9, 'Product 9', 'Brand Z', 'Vendor A'),
+(10, 'Product 10', 'Brand X', 'Vendor B');
+")
+
+# Step 2: Create the 'raw' table
+dbExecute(con, "
+CREATE TABLE raw (
+    nowtime DATETIME,
+    product_id INTEGER,
+    current_price REAL,
+    old_price REAL,
+    price_per_unit REAL,
+    other TEXT
+);
+")
+
+# Insert data into the 'raw' table using a recursive query
+dbExecute(con, "
+WITH RECURSIVE series(x) AS (
+    SELECT 1
+    UNION ALL
+    SELECT x + 1
+    FROM series
+    WHERE x <= 1000
+)
+INSERT INTO raw (nowtime, product_id, current_price, old_price, price_per_unit, other)
+SELECT 
+    DATETIME('2024-01-01 00:00:00', '+' || (x - 1) || ' hours') AS nowtime,
+    ABS(RANDOM() % 10) + 1 AS product_id,
+    ROUND(RANDOM() % 100 + 1, 2) AS current_price,
+    ROUND(RANDOM() % 100 + 1, 2) AS old_price,
+    ROUND(RANDOM() % 10 + 0.1, 2) AS price_per_unit,
+    CASE (RANDOM() % 3)
[…]
